Concordance between observations and simulations in the evolution of the mass relation between supermassive black holes and their host galaxies

### 2022-04-28 projects/2020_HSC_to_numerical_simulation/manuscript/manuscript.tex
Short the abs

\begin{abstract}
We carry out a comparative analysis of the relation between the mass of supermassive black holes (BHs) and the stellar mass of their host galaxies (\mbh --\smass) at $0.2<z<1.7$ using well-matched observations and multiple state-of-the-art simulations (e.g., Massive Black II, Horizon-AGN, Illustris, TNG and a semi-analytic model). The observed sample consist of 646 uniformly-selected SDSS quasars ($0.2 < z < 0.8$) and 32 broad-line AGNs ($1.2<z<1.7$) with imaging from Hyper Suprime-Cam (HSC) for the former and Hubble Space Telescope (HST) for the latter. To make fair comparisons, we first add realistic observational uncertainties to the simulation data and then construct a simulated sample in the same manner as the observations. Over the full redshift range, our analysis demonstrates that all simulations predict a level of intrinsic scatter of the scaling relations comparable to the observations which appear to agree with the dispersion of the local relation. 
Regarding the mean scaling relation, Horizon-AGN and TNG are in closest agreement with the observational data at both low and high redshift ($z\sim$ 0.2 and 1.5, respectively) while the other simulations show subtle differences within the uncertainties. 
For insight into the physics involved, the scatter of the scaling relation, seen in the SAM, is reduced by a factor of two and thus closer to the observations after adopting a new feedback model which considers the geometry of the AGN outflow. The consistency in the dispersion with redshift in our analysis lends support to the importance of both quasar and radio mode feedback prescriptions in the simulations. Finally, we investigate evolution as a function of stellar and \red{BH} mass, and argue that the most direct way to probe the co-evolution of galaxies and \red{BH}s is to increase the sensitivity, e.g., with JWST, thereby pushing to lower masses and minimizing possible biases arising from selection effects.

\end{abstract}
\keywords{Galaxy evolution (594); Active galaxies (17); Active galactic nuclei (16)}

@@

\subsection{Horizon-AGN}\label{subsec:Horizon}
The simulation Horizon-AGN~\citep{2014MNRAS.444.1453D, 2016MNRAS.463.3948D} has a volume of 142 cMpc$^3$ and was generated using the adaptive mesh refinement code {\tt Ramses}~\citep{2002A&A...385..337T} with a $\Lambda$CDM model based on WMAP7~\citep{2011ApJS..192...18K} cosmological results. The dark matter particle mass is $8\times 10^7 M_{\odot}$. The stellar particle mass is $2\times 10^6 M_{\odot}$ and the MBH seed mass is $10^5 M_{\odot}$. Adaptive mesh refinement is permitted down to $\Delta x=1$~kpc, and, if the total mass in a cell becomes greater than 8 times the initial mass resolution, it is performed in a quasi-Lagrangian manner. Collisionless particles (dark matter and star particles) are evolved using a particle-mesh solver with a cloud-in-cell interpolation.

%The simulation includes gas cooling down to $10^4\, \rm K$ \citep{sutherland&dopita93}, and stochastic star formation with a constant star formation efficiency $\epsilon_*=0.02$, which occurs in regions where the gas number density exceeds the star formation threshold $n_0 = 0.1\,\rm H\, cm^{-3}$. %A Salpeter IMF is assumed. 
%\red{Stellar feedback} is modeled as mechanical energy injection from Type Ia SNe, Type II SNe and stellar winds, with the metal enrichment from these sources.

@@

For dark matter halos that merge with a larger halo, the impact of dynamical friction is assessed to define whether the halo will survive as a satellite or sink to the center of the dominant galaxy which increases its mass. The binary interactions (fly-bys and mergers), among satellite sub-halos, are  also described by the model. In each halo, we compute the fraction of gas which cools because of the atomic processes and settles into a disk~\citep{Mo1998}. The stars are converted from the gas through three channels: (1) quiescent star formation with long time scales: $\sim1$~Gyr; (2) starbursts following galaxy interactions with timescales $\lesssim 100$~Myr, according to BH feeding; \todo{Nicola: the link between BH feeding and star formation channels in this sentence is not clear.} (3) the loss of angular momentum triggered by the internal disk instabilities causing the gas inflows to the center, resulting in stimulated star formation (as well as BH accretion). The stellar feedback is also considered by calculating the energy released by the supernovae associated with the total star formation which returns a fraction of the disk gas into a hot phase. %A Salpeter IMF is adopted in the SAM simulation. 
\red{A $\Lambda$CDM power} spectrum of perturbations with a total matter density parameter $\Omega_0=0.3$, a baryon density parameter 
$\Omega_b=0.04$, a dark energy density parameter $\Omega_\Lambda=0.7$, and a Hubble constant $h$=0.7 is adopted.

@@

The BH accretion is based on both interaction-driven and disk instability feeding modes.
The SAM adopted here implements a new and improved  model for the AGN feedback with respect to the previous versions~\citep{Menci2008}. In both versions, the basic assumption is that fast winds with velocity up to $10^{-1}c$ observed in the central regions of AGNs~\citep{Chartas2002, Pounds2003}  result in  supersonic outflows that compress the gas into a blast wave terminated by a leading shock front. This  moves outward with a lower but still supersonic speed, and sweeps out the surrounding medium. However, while in the earlier version of the SAM~\citep{Menci2016} the blast wave is assumed to expand into an isotropically distributed medium, in the new description of AGN feedback~\citep{Menci2019} the full two-dimensional structure of the gas disc and of the expanding blast wave is followed in detail. The main physical difference is that in the new model the large density of gas along the plane of the disc causes the blast wave expansion to stall in such a direction, while it expands with large velocities in the vertical direction. The resulting strong dependence of the total (integrated over directions) outflow rate on the AGN luminosity $L_{AGN}$ and on the gas content of the galaxy $M_{gas}$ is shown in Figure~\ref{fig:SAM}. Such a new AGN feedback model has been tested in detail against a state-of-the-art compilation of observed outflows in 19 galaxies with different measured gas and dynamical masses~\citep{Fiore2017}, allowing for a detailed, one-by-one comparison with the model predictions. This well tested AGN feedback model allowed us to derive, for each simulated galaxy in the SAM,  the outflow expansion and the mass outflow rates in different directions with respect to the plane of the disc.

>>>>

\begin{abstract}
We carry out a comparative analysis of the relation between the mass of supermassive black holes (BHs) and the stellar mass of their host \red{galaxies} at $0.2<z<1.7$ using well-matched observations and multiple state-of-the-art simulations (e.g., Massive Black II, Horizon-AGN, Illustris, TNG and a semi-analytic model). The observed sample consist of 646 uniformly-selected SDSS quasars ($0.2 < z < 0.8$) and 32 broad-line AGNs ($1.2<z<1.7$) with imaging from Hyper Suprime-Cam (HSC) for the former and Hubble Space Telescope (HST) for the latter. \red{We} first add realistic observational uncertainties to the simulation data and then construct a simulated sample in the same manner as the observations. Over the full redshift range, our analysis demonstrates that all simulations predict a level of intrinsic scatter of the scaling relations comparable to the observations which appear to agree with the dispersion of the local relation. Regarding the mean relation, Horizon-AGN and TNG are in closest agreement with the \red{observations} at low and high redshift ($z\sim$ 0.2 and 1.5, respectively) while the other simulations show subtle differences within the uncertainties. For insight into the physics involved, the scatter of the scaling relation, seen in the SAM, is reduced by a factor of two and closer to the observations after adopting a new feedback model which considers the geometry of the AGN outflow. The consistency in the dispersion with redshift in our analysis \red{supports} the importance of both quasar and radio mode feedback prescriptions in the simulations. Finally, we \red{highlight the importance of increasing} the sensitivity, e.g., JWST, thereby pushing to lower masses and minimizing biases \red{due to} selection effects.
\end{abstract}
\keywords{Galaxy evolution (594); Active galaxies (17); Active galactic nuclei (16)}

@@

\subsection{Horizon-AGN}\label{subsec:Horizon}
The simulation Horizon-AGN~\citep{2014MNRAS.444.1453D, 2016MNRAS.463.3948D} has a volume of 142 cMpc$^3$ and was generated using the adaptive mesh refinement code {\tt Ramses}~\citep{2002A&A...385..337T} with a $\Lambda$CDM model based on WMAP7~\citep{2011ApJS..192...18K} cosmological results. The dark matter particle mass is $8\times 10^7 M_{\odot}$. The stellar particle mass is $2\times 10^6 M_{\odot}$ and the MBH seed mass is $10^5 M_{\odot}$. Adaptive mesh refinement is permitted down to $\Delta x=1$~kpc, and, if the total mass in a cell becomes greater than 8 times the initial mass resolution, it is performed in a quasi-Lagrangian manner. Collisionless particles (dark matter and star particles) are evolved using a particle-mesh solver with a cloud-in-cell interpolation.
\red{The simulation includes gas cooling down to $10^4\, \rm K$ \citep{sutherland&dopita93}, and stochastic star formation. Stellar feedback is modeled as mechanical energy injection from Type Ia SNe, Type II SNe and stellar winds, with the metal enrichment from these sources.}
%The simulation includes gas cooling down to $10^4\, \rm K$ \citep{sutherland&dopita93}, and stochastic star formation with a constant star formation efficiency $\epsilon_*=0.02$, which occurs in regions where the gas number density exceeds the star formation threshold $n_0 = 0.1\,\rm H\, cm^{-3}$. %A Salpeter IMF is assumed. 
%\red{Stellar feedback} is modeled as mechanical energy injection from Type Ia SNe, Type II SNe and stellar winds, with the metal enrichment from these sources.

@@

For dark matter halos that merge with a larger halo, the impact of dynamical friction is assessed to define whether the halo will survive as a satellite or sink to the center of the dominant galaxy which increases its mass. The binary interactions (fly-bys and mergers), among satellite sub-halos, are  also described by the model. In each halo, we compute the fraction of gas which cools because of the atomic processes and settles into a disk~\citep{Mo1998}. The stars are converted from the gas through three channels: (1) quiescent star formation with long time scales: $\sim1$~Gyr; (2) starbursts following galaxy interactions with timescales $\lesssim 100$~Myr;  (3) the loss of angular momentum triggered by the internal disk instabilities causing the gas inflows to the center, resulting in stimulated star formation (as well as BH accretion). The stellar feedback is also considered by calculating the energy released by the supernovae associated with the total star formation which returns a fraction of the disk gas into a hot phase. %A Salpeter IMF is adopted in the SAM simulation. 
\red{A $\Lambda$CDM power} spectrum of perturbations with a total matter density parameter $\Omega_0=0.3$, a baryon density parameter 
$\Omega_b=0.04$, a dark energy density parameter $\Omega_\Lambda=0.7$, and a Hubble constant $h$=0.7 is adopted.

@@

\red{The BH accretion is assumed to follow from the gas instabilities resulting from either galaxy interactions and disk instabilities, and is thus related to star formations channels 2-3: a fraction of the cold gas destabilized during galaxy encounters and through disc instabilities is in fact accreted onto the central BHs (with the remaining fraction fueling star formation through channels 2-3 described above). Such fractions and the corresponding timescales for accretion are computed as described in~\citet[][in Section 3]{Menci2014}.}

The SAM adopted here implements a new and improved  model for the AGN feedback with respect to the previous versions~\citep{Menci2008}. In both versions, the basic assumption is that fast winds with velocity up to $10^{-1}c$ observed in the central regions of AGNs~\citep{Chartas2002, Pounds2003}  result in  supersonic outflows that compress the gas into a blast wave terminated by a leading shock front. This  moves outward with a lower but still supersonic speed, and sweeps out the surrounding medium. However, while in the earlier version of the SAM~\citep{Menci2016} the blast wave is assumed to expand into an isotropically distributed medium, in the new description of AGN feedback~\citep{Menci2019} the full two-dimensional structure of the gas disc and of the expanding blast wave is followed in detail. The main physical difference is that in the new model the large density of gas along the plane of the disc causes the blast wave expansion to stall in such a direction, while it expands with large velocities in the vertical direction. The resulting strong dependence of the total (integrated over directions) outflow rate on the AGN luminosity $L_{AGN}$ and on the gas content of the galaxy $M_{gas}$ is shown in Figure~\ref{fig:SAM}. Such a new AGN feedback model has been tested in detail against a state-of-the-art compilation of observed outflows in 19 galaxies with different measured gas and dynamical masses~\citep{Fiore2017}, allowing for a detailed, one-by-one comparison with the model predictions. This well tested AGN feedback model allowed us to derive, for each simulated galaxy in the SAM,  the outflow expansion and the mass outflow rates in different directions with respect to the plane of the disc.
